High-Performance Primitives for Distributed Memory Architectures using Asynchronous Events

### 2012-11-12 doc/llr-pldi2013/interface.tex
Minor edits to the low level paper

A {\tt ptr\_t} value returned by the {\tt alloc}
method may be used to access the corresponding element in any of the instances of the region (line 62).
This
allows the creation of irregular data structures within physical regions.  The {\tt ptr\_t} values can
be safely stored in a region, and need not be changed when data is copied from one instance to another.
In common cases, this is as cheap as array-addressing calculations, so the overhead is minimal.
When more than one instance exists for a region, the client is responsible for maintaining coherence
of the data between the regions.  This is typically achieved through the use of instance-to-instance
copy operations (line 63) that use the producer task's completion event as a prerequisite and provide
their own completion event as a prerequisite for the consumer task.  When a copy operation is
the last use of a particular instance, the destruction of that instance can also be requested and made
dependent on the copy operation's event, allowing the runtime to reclaim the memory as soon as (but no
sooner than) possible.

If the only operations a task will perform on a instance are reductions, a special reduction-only
instance can be created (line 55).  Rather than applying each reduction separately to the target instance,
the reduction operations can be accumulated in the reduction instance.  Using the {\tt reduce\_to} 
method (line 64), these accumulated reductions can be applied in bulk to another instance, which may be either
a normal instance or another reduction-only instance.  First, it results
in much better performance, as we show in Section~\ref{subsec:reducmicro}, by reducing both
the total inter-node traffic and the overhead of that traffic while still allowing parallelism.
Second, a reduction-only instance is often smaller than a normal instance, potentially improving
cache performance and helping address capacity issues for smaller memories (e.g. a GPU's device
memory).  For example, if reductions are being made to a single field in a larger structure, the 
reduction-only instance only needs to store that field and not the other fields that won't be 
modified by the reduction operation.

>>>>

A {\tt ptr\_t} value returned by the {\tt alloc}
method may be used to access the corresponding element in any of the instances of the region (line 62).
This
allows the creation of irregular data structures within physical regions.  The {\tt ptr\_t} values can
be safely stored in a region, and need not be changed when data is copied from one instance to another.
In common cases, this is as cheap as array-addressing calculations, so the overhead is minimal.
When more than one instance exists for a region, the client is responsible for maintaining coherence
of the data between instances of a region.  This is typically achieved through the use of instance-to-instance
copy operations (line 63) that use the producer task's completion event as a prerequisite and provide
their own completion event as a prerequisite for the consumer task.  When a copy operation is
the last use of a particular instance, the destruction of that instance can also be requested and made
dependent on the copy operation's event (line 56), allowing the runtime to reclaim the memory as soon as (but no
sooner than) possible.

If the only operations a task will perform on a instance are reductions, a special reduction-only
instance can be created (line 55).  Rather than applying each reduction separately to the target instance,
the reduction operations can be accumulated in the reduction instance.  Using the {\tt reduce\_to} 
method (line 64), these accumulated reductions can be applied in bulk to another instance, which may be either
a normal instance or another reduction-only instance.  The use of reduction instances can result 
in much better performance, as we show in Section~\ref{subsec:reducmicro}, by reducing both
the total inter-node traffic and the overhead of that traffic while still allowing parallelism.
Furthermore, reduction-only instances are often smaller than normal instances, potentially improving
cache performance and helping address capacity issues for smaller memories (e.g. a GPU's device
memory).  For example, if reductions are being made to a single field in a larger structure, the 
reduction-only instance only needs to store that field and not the other fields that won't be 
modified by the reduction operation.

### 2012-11-12 doc/llr-pldi2013/micro.tex
Minor edits to the low level paper

For our experiment, there are 8 reduction tasks created for each node, one per processor.  Each task performs 4M reductions
into a histogram with 256K buckets.  The input data is random, so although the average number of reductions per bucket is 16, there will be some buckets with more and some with fewer.  The results for a variety of node counts are shown in Figure~\ref{fig:reducbench}.
They show that the version using reduction-only instance performs 
best and scales nearly linearly with the number of nodes, achieving over a billion reductions
per second in the 16 node case.  The localized instance approach works well for a single node, 
but its inherently serial nature doesn't benefit from increasing node counts.  Finally, the version using individual
reduction messages is scalable, but the need to send each reduction separately and the per-message overhead make it nearly 1000x slower than the reduction instance version that is
able to accumulate multiple reductions to the same location locally and move data in bulk.

>>>>

For our experiment, there are 8 reduction tasks created for each node, one per processor.  Each task performs 4M reductions
into a histogram with 256K buckets.  The input data is random, so although the average number of reductions per bucket is 16, there will be some buckets with more and some with fewer.  The results for a variety of node counts are shown in Figure~\ref{fig:reducbench}.
They show that the version using reduction-only instance performs 
best and scales nearly linearly with the number of nodes, achieving over a billion reductions
per second in the 16 node case.  The localized instance approach works well for a single node, 
but its inherently serial nature doesn't benefit from increasing node counts.  Finally, the version using individual
reduction messages is scalable, but the need to send each reduction separately and the per-message overhead make it nearly 
1000x slower than the reduction instance version, which is
able to accumulate multiple reductions to the same location locally and move data in bulk.
